---
title: "Regression Project (Life Expectancy) (Code)"
subtitle: Math 4387/5387 Applied Regression Analysis
format:
  html:
    code-tools: true
    self-contained: true
author: Diego Garza Valdez
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis

```{r, include = FALSE}

library(dplyr)
library(ggplot2)
library(leaps)
library(caret)
library(car)
library(lmtest)
library(api2lm)
library(alr4)
library(effects)

```

```{r}
life_exp_raw<- read.csv("https://raw.githubusercontent.com/Diego150103/Regression/refs/heads/main/Life_Expectancy.csv")
```

```{r}
life_exp_clean <- 
  life_exp_raw |>
  subset(Year == 2015,select = c("Country", "Life.expectancy", "Adult.Mortality",
                                 "BMI", "GDP", "Schooling", "HIV.AIDS",
                                 "Population", "Status"))
  
```

```{r}
summary(life_exp_clean)
```

```{r}
life_exp_clean <-
  life_exp_clean |>
  transform(Status = factor(Status))
```

```{r}
summary(life_exp_clean)
```

```{r}
names(life_exp_clean) <- c("Country",
                           "life_expectancy",
                           "adult_mortality",
                           "BMI",
                           "GDP",
                           "Schooling",
                           "HIV_AIDS", 
                           "Population",
                           "Status")
```

```{r}
summary(life_exp_clean)
```

### Final Selected Variables

| Variable | Description | Type |
|----|----|----|
| `life_expectancy` | Average life expectancy at birth (years) | Numeric |
| `adult_Mortality` | Adult mortality rate per 1000 individuals | Numeric |
| `BMI` | Average Body Mass Index | Numeric |
| `GDP` | Gross Domestic Product per capita (USD per year) | Numeric |
| `Schooling` | Average number of years of schooling | Numeric |
| `HIV_AIDS` | Prevalence of HIV/AIDS (% of population) | Numeric |
| `Population` | Total population of the country | Numeric |
| `Status` | Country classification as Developed or Developing | Factor |

### **Rregressor graphs**

### **Adult mortality**

```{r}
ggplot(life_exp_clean, aes(x = adult_mortality)) +
  geom_histogram(binwidth = 25, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Adult Mortality", x = "Adult Mortality (per 1000)", y = "Frequency")

```

### Adult mortality vs. Life Expectancy

```{r}
ggplot(life_exp_clean, aes(x = adult_mortality, y = life_expectancy)) +
  geom_point(color = "darkblue") +
  labs(title = "Life Expectancy by Adult Mortality Rate", x = "Adult Mortality", 
       y = "Life Expectancy")
```

### BMI

```{r}
# Histogram for BMI
ggplot(life_exp_clean, aes(x = BMI)) +
  geom_histogram(binwidth = 5, fill = "lightgreen", color = "black") +
  labs(title = "Distribution of BMI", x = "BMI", y = "Frequency")

```

### BMI **vs. Life Expectancy**

```{r}
# Scatter Plot for BMI vs. Life Expectancy
ggplot(life_exp_clean, aes(x = BMI, y = life_expectancy)) +
  geom_point(color = "red") +
  labs(title = "Life Expectancy by BMI", x = "BMI", y = "Life Expectancy")
```

### GDP

```{r}
# Histogram for GDP per Capita
ggplot(life_exp_clean, aes(x = GDP)) +
  geom_histogram(binwidth = 5000, fill = "purple", color = "black") +
  labs(title = "Distribution of GDP per Capita", x = "GDP per Capita (USD)", y = "Frequency")
```

**log-transformation for GDP**

```{r}
# Apply a log transformation to GDP
life_exp_clean <- 
  life_exp_clean |>
  mutate(log_GDP = log(GDP))


summary(life_exp_clean)
```

### log_GDP

```{r}
# Histogram for GDP per Capita
ggplot(life_exp_clean, aes(x = log_GDP)) +
  geom_histogram(fill = "purple", color = "black") +
  labs(title = "Distribution of log GDP per Capita", x = "log GDP per Capita (log USD)", y = "Frequency")
```

### **log_GDP vs. Life Expectancy**

```{r}
# Scatter Plot for GDP vs. Life Expectancy
ggplot(life_exp_clean, aes(x = log_GDP, y = life_expectancy)) +
  geom_point(color = "purple") +
  labs(title = "Life Expectancy by log_GDP per Capita", x = "log_GDP per Capita", 
       y = "Life Expectancy")
```

### Schooling

```{r}
# Histogram for Schooling
ggplot(life_exp_clean, aes(x = Schooling)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "black") +
  labs(title = "Distribution of Years of Schooling", x = "Years of Schooling", y = "Frequency")

```

### **Schooling vs. Life Expectancy**

```{r}
# Scatter Plot for Schooling vs. Life Expectancy
ggplot(life_exp_clean, aes(x = Schooling, y = life_expectancy)) +
  geom_point(color = "black") +
  labs(title = "Life Expectancy by Years of Schooling", x = "Years of Schooling", 
       y = "Life Expectancy")
```

### HIV_AIDS

```{r}
# Histogram for HIV/AIDS Prevalence
ggplot(life_exp_clean, aes(x = HIV_AIDS)) +
  geom_histogram(binwidth = 1, fill = "orange", color = "black") +
  labs(title = "Distribution of HIV/AIDS Prevalence", x = "HIV/AIDS Prevalence (%)", y = "Frequency")

```

### **Make HIV_AIDS categorical**

```{r}
# Categorize HIV_AIDS prevalence
life_exp_clean <- life_exp_clean |>
  mutate(
    HIV_AIDS_level = case_when(
      HIV_AIDS <= 0.4 ~ "Low",           # Prevalence <= 0.4%
      HIV_AIDS > 0.4 & HIV_AIDS <= 5 ~ "Moderate", # Prevalence > 0.4% & <= 5%
      HIV_AIDS > 5 ~ "High"             # Prevalence > 5%
    )
  )

# Declare levels for the new factor variable
life_exp_clean$HIV_AIDS_level <- factor(
  life_exp_clean$HIV_AIDS_level,
  levels = c("Low", "Moderate", "High")
)
```

```{r}
summary(life_exp_clean$HIV_AIDS_level)
```

### Box plot HIV_AIDS

```{r}
# Bar Chart for HIV_AIDS Categories
ggplot(life_exp_clean, aes(x = HIV_AIDS_level, fill = HIV_AIDS_level)) +
  geom_bar(color = "black") +
  labs(
    title = "Distribution of HIV/AIDS Prevalence Categories",
    x = "HIV/AIDS Prevalence Category",
    y = "Count"
  ) +
  scale_fill_manual(values = c("green", "orange", "red")) 

```

### **HIV_AIDS vs. Life Expectancy**

```{r}
# Box Plot for HIV_AIDS vs Life Expectancy
ggplot(life_exp_clean, aes(x = HIV_AIDS_level, y = life_expectancy, fill = HIV_AIDS_level)) +
  geom_boxplot(color = "black") +
  labs(
    title = "Life Expectancy by HIV/AIDS Prevalence Category",
    x = "HIV/AIDS Prevalence Category",
    y = "Life Expectancy"
  ) +
  scale_fill_manual(values = c("green", "orange", "red"))

```

### Population

```{r}
# Histogram for Population
ggplot(life_exp_clean, aes(x = Population)) +
  geom_histogram( fill = "blue", color = "black") +
  labs(
    title = "Distribution of Population",
    x = "Population",
    y = "Frequency"
  ) 

```

```{r}
summary(life_exp_clean$Population)
```

```{r}
# Apply log transformation to Population
life_exp_clean <- life_exp_clean |>
  mutate(log_Population = log(Population + 1))
```

```{r}
# Histogram for Log-Transformed Population
ggplot(life_exp_clean, aes(x = log_Population)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  labs(
    title = "Distribution of Log-Transformed Population",
    x = "Log(Population)",
    y = "Frequency"
  ) 
```

### log Population vs Life Expectancy

```{r}
# Scatter Plot for Log-Transformed Population vs Life Expectancy
ggplot(life_exp_clean, aes(x = log_Population, y = life_expectancy)) +
  geom_point(color = "blue", alpha = 0.7) +
  labs(
    title = "Life Expectancy vs Log-Transformed Population",
    x = "Log(Population)",
    y = "Life Expectancy"
  ) 

```

### Status

```{r}
# Bar Plot for Country Status
ggplot(life_exp_clean, aes(x = Status, fill = Status)) +
  geom_bar(color = "black") +
  labs(title = "Distribution of Country Status", x = "Country Status", y = "Count") +
  scale_fill_manual(values = c("red", "cyan"))

```

### **Status vs. Life Expectancy**

```{r}
# Boxplot for Status vs. Life Expectancy
ggplot(life_exp_clean, aes(x = Status, y = life_expectancy, fill = Status)) +
  geom_boxplot() +
  labs(title = "Life Expectancy by Country Status", x = "Country Status", 
       y = "Life Expectancy")
```

```{r}
# Subset the data frame to include the necessary variables
final_life_exp <- life_exp_clean |>
  select(
    life_expectancy,
    adult_mortality,
    BMI,
    log_GDP,
    Schooling,
    HIV_AIDS_level,         
    log_Population,   
    Status            
  )
```

```{r}
# Create a new copy of the data frame to reset indices
final_life_exp <- as.data.frame(final_life_exp)

# Remove any row names
rownames(final_life_exp) <- NULL
```

```{r}
summary(final_life_exp)
```

## Variable Selection

```{r}
complete_lmod <- lm(life_expectancy ~ ., data = final_life_exp)
```

```{r}
summary(complete_lmod)
```

### **Best subset regression with** ${R_{a}^2}$

```{r}
rs <- leaps::regsubsets(formula(complete_lmod), data = final_life_exp)
```

```{r}
plot(rs, scale = "adjr2")
```

```{r}
# Fit the regression model
best_subset_lmod <- lm(life_expectancy ~ adult_mortality + 
                    Schooling + HIV_AIDS_level + log_Population + 
                    Status, data = final_life_exp)

# Display the summary of the model
summary(best_subset_lmod)
```

#### **Model formula (Best subset)**

$$
\begin{aligned}&E(\mathtt{life\_expectancy} \mid \mathtt{adult\_mortality}, \mathtt{Schooling}, \mathtt{HIV\_AIDS\_level}, \mathtt{log\_Population}, \mathtt{Status}) \\&\qquad = 62.294 - 0.026 \mathtt{adult\_mortality} + 1.264 \mathtt{Schooling} - 5.412 \mathtt{HIV\_AIDS_{Moderate}} - 3.407 \mathtt{HIV\_AIDS_{High}} \\&\qquad \quad - 0.017 \mathtt{log\_Population} - 1.986 \mathtt{Status_{Developing}}.\end{aligned}
$$

### **Stepwise selection**

```{r}
# Remove rows with missing values
clean_life_exp <- na.omit(final_life_exp)

# Fit the complete model
complete_lmod_2 <- lm(life_expectancy ~ ., data = clean_life_exp)

# Perform stepwise selection
stepwise_reg <- step(complete_lmod_2, direction = "both")
```

```{r}
step_lmod <- lm(life_expectancy ~ adult_mortality + Schooling + HIV_AIDS_level + Status, 
                 data = clean_life_exp)

summary(step_lmod)

```

#### **Model formula (Stepwise)**

$$
\begin{aligned}
&E(\mathtt{life\_expectancy} \mid \mathtt{adult\_mortality}, \mathtt{Schooling}, \mathtt{HIV\_AIDS\_level}, \mathtt{Status}) \\
&\qquad = 61.089 - 0.027 \mathtt{adult\_mortality} + 1.326 \mathtt{Schooling} \\
&\qquad \quad - 5.362 \mathtt{HIV\_AIDS_{Moderate}} - 3.016 \mathtt{HIV\_AIDS_{High}} - 1.747 \mathtt{Status_{Developing}}.
\end{aligned}
$$

### Mallow’s $C_p$ Statistic

```{r}
srs <- summary(rs)
p <- seq_along(srs$cp) + 1
plot(srs$cp ~ p, ylab = "Mallow's Cp", ylim = c(0, max(p)))
abline(0, 1)
```

```{r}
srs$which
```

### The best model

```{r}
final_best_lmod <- lm(life_expectancy ~ adult_mortality + Schooling + 
                        HIV_AIDS_level + log_Population + Status, 
                      data = final_life_exp)
```

```{r}
coef(final_best_lmod)
```

$$
\begin{aligned}&E(\mathtt{life\_expectancy} \mid \mathtt{adult\_mortality}, \mathtt{Schooling}, \mathtt{HIV\_AIDS\_level}, \mathtt{log\_Population}, \mathtt{Status}) \\&\qquad = 62.294 - 0.026 \mathtt{adult\_mortality} + 1.264 \mathtt{Schooling} - 5.412 \mathtt{HIV\_AIDS_{Moderate}} - 3.407 \mathtt{HIV\_AIDS_{High}} \\&\qquad \quad - 0.017 \mathtt{log\_Population} - 1.986 \mathtt{Status_{Developing}}.\end{aligned}
$$

## Checking assumptions

### Structural assumptions

```{r}
residualPlot(final_best_lmod, quadratic = FALSE)
```

```{r}
residualPlots(final_best_lmod, 
              tests = FALSE, quadratic = FALSE, fitted = FALSE)
```

### Checking for influential observations

#### **Outliers**

```{r}
outlier_plot(final_best_lmod, id_n = 2)
```

```{r}
outlier_test(final_best_lmod)
```

#### Leverage points

```{r}
leverage_plot(final_best_lmod)
```

```{r}
(leverage_countries <- life_exp_clean[c(93, 156, 183), ])
```

#### Influential observations

```{r}
cooks_plot(final_best_lmod, id_n = 3)
```

```{r}
 final_life_exp[c(93, 182, 183), ]
```

```{r}
(influencial_countries <- life_exp_clean[c(93, 182, 183), ])
```

### **Constant Variance**

```{r}
residualPlot(final_best_lmod, type = "rstandard", quadratic = FALSE)
```

```{r}
plot(final_best_lmod, which = 3)
```

### **Checking normality of errors**

```{r}
qqPlot(final_best_lmod)
```

```{r}
shapiro.test(rstandard(final_best_lmod))
```

## Inferences

### Comparing complete model against final model

So that we don't get this problem ("models were not all fitted to the same size of dataset") we do the following

```{r}
complete_data <- final_life_exp[complete.cases(final_life_exp[, c("adult_mortality", "BMI", "log_GDP", 
                                                                  "Schooling", "HIV_AIDS_level", 
                                                                  "log_Population", "Status")]), ]
```

```{r}
complete_lmod_refit <- lm(life_expectancy ~ ., data = complete_data)

final_best_lmod_refit <- lm(life_expectancy ~ adult_mortality + Schooling + 
                             HIV_AIDS_level + log_Population + Status, 
                             data = complete_data)
```

$$
\begin{aligned}
&H_0: \beta_{\mathtt{BMI}} = \beta_{\mathtt{log\_GDP}} = 0 \mid \{\beta_{\mathtt{Intercept}}, \beta_{\mathtt{adult\_mortality}}, \beta_{\mathtt{Schooling}}, \beta_{\mathtt{HIV\_AIDS\_level}}, \beta_{\mathtt{log\_Population}}, \beta_{\mathtt{Status}}\} \in \mathbb{R}^6.\\
&H_a: \beta_{\mathtt{BMI}} \neq 0 \text{ or } \beta_{\mathtt{log\_GDP}} \neq 0 \mid \{\beta_{\mathtt{Intercept}}, \beta_{\mathtt{adult\_mortality}}, \beta_{\mathtt{Schooling}}, \beta_{\mathtt{HIV\_AIDS\_level}}, \beta_{\mathtt{log\_Population}}, \beta_{\mathtt{Status}}\} \in \mathbb{R}^6.
\end{aligned}
$$

```{r}
anova(final_best_lmod_refit, complete_lmod_refit)
```

The ANOVA results show no significant difference between the two models `p-value = 0.72`. This indicates that including additional predictors like `BMI` and `log_GDP` in the complete model does not significantly improve the fit compared to the final best model. Therefore, the simpler final best model is preferable.

### **Confidence intervals for parameters**

```{r}
confint_adjust(final_best_lmod, method = "bonferroni")
```

```{r}
confint_adjust(final_best_lmod, method = "wh")
```

## Collinearity

```{r}
vif(final_best_lmod)
```

We can see that the VIF values for all the variables in our model are below the threshold of 10, thus there are no collinearity issues in our model.

## Effect plots

```{r}
plot(allEffects(final_best_lmod))
```
